---
title: "  model for Theoretical Limit"
author: "Jonathan Gill"
date: "April 24, 2018"
output: html_document
---

Does not seem promising. Tried rpart() on data with NAs omitted, and on data with NAs converted to means (not sure what that really means here). For NA omission, I did not alter the NAs in the test set. I tried messing around with cp cuttoffs. Couldn't really get the evaluation metric below 0.89.

I think this may be because there is a discrete number of probability levels generated. Since the evaluation metric scales with log of the probability estimates, it might be better to have something continuous. This is just a guess.

split up data into train and test
```{r}
set.seed = 1
samp = sample.int(n = nrow(train), size = 3000, replace = F)
testsub = train[samp, ]
trainsub = train[-samp,]
```



with NAs omitted:

```{r}

train = read.csv("train.csv")

#Converting relevant variables to factors
cols1 = c(2,3,4,5,6,8,11,12,14:45,48)
train[cols1] = lapply(train[cols1], factor)

#removing aliased columns
train = train[,-which(names(train) %in% c("density_rural","single"))]


#split dataset here if desired


set.seed(1)
require(rpart)
tree = rpart(suppdem ~ . , data = trainsub, na.action = na.omit, method = "class", cp = 0.0015)
plotcp(tree)
pruned_tree = prune(tree, cp=0.0043)
plotcp(pruned_tree)
require(rpart.plot)
prp(pruned_tree, extra=100)




#testing
predictions = predict(pruned_tree, testsub ,type = "prob")
#just want the probability of "Yes"
predictions = predictions[,1] 
#change response from factor to binary
test_binary = ifelse(testsub$suppdem=="Y", 1, 0)
d = -(1/nrow(testsub))*sum((test_binary *log(predictions) + (1-test_binary )*log(1-predictions)))
```



With NAs converted:
```{r}
na.convert.mean = function (frame) 
{
    vars <- names(frame)
    if (!is.null(resp <- attr(attr(frame, "terms"), "response"))) {
        vars <- vars[-resp]
        x <- frame[[resp]]
        pos <- is.na(x)
        if (any(pos)) {
            frame <- frame[!pos, , drop = FALSE]
            warning(paste(sum(pos), "observations omitted due to missing values in the response"))
        }
    }
    for (j in vars) {  #j is variable names
        x <- frame[[j]]
        pos <- is.na(x)
        if (any(pos)) {
            if (length(levels(x))) {   # factors
                xx <- as.character(x)
                xx[pos] <- "NA"
                x <- factor(xx, exclude = NULL)
            }
            else if (is.matrix(x)) {   # matrices
                ats <- attributes(x)
                x.na <- 1*pos
#               x[pos] <- 0
                w <- !pos
                n <- nrow(x)
                TT <- array(1, c(1, n))
                xbar <- (TT %*% x)/(TT %*% w)
                xbar <- t(TT) %*% xbar
                x[pos] <- xbar[pos]
                attributes(x) <- ats
                attributes(x.na) <- ats
                dimnames(x.na)[[2]]=paste(dimnames(x)[[2]],".na",sep='')
                frame[[paste(j,".na",sep='')]] <- x.na 
            } else {   # ordinary numerical vector
                ats <- attributes(x)
                x[pos] <- mean(x[!pos])
#               x[pos] <- 0
                x.na <- 1*pos
                frame[[paste(j,".na",sep='')]] <- x.na 
                attributes(x) <- ats
            }
            frame[[j]] <- x
        }
    }
    frame
}




train = read.csv("train.csv")

train_convert_na = na.convert.mean(train)

#Converting relevant variables to factors
cols1 = c(2,3,4,5,6,8,11,12,14:45,48)
train_convert_na[cols1] = lapply(train_convert_na[cols1], factor)

#removing aliased columns
train_convert_na = train_convert_na[,-which(names(train_convert_na) %in% c("density_rural","single"))]



#split dataset here if desired

set.seed(1)
require(rpart)
tree = rpart(suppdem ~ . , data = trainsub, method = "class", cp = 0.0015)
plotcp(tree)
pruned_tree = prune(tree, cp=0.034)
plotcp(pruned_tree)
require(rpart.plot)
prp(pruned_tree, extra=100)


#testing
predictions = predict(pruned_tree, testsub ,type = "prob")
#just want the probability of "Yes"
predictions = predictions[,1] 
#change response from factor to binary
test_binary = ifelse(testsub$suppdem=="Y", 1, 0)
d = -(1/nrow(testsub))*sum((test_binary *log(predictions) + (1-test_binary )*log(1-predictions)))
d
```


Testing
```{r}

predictions = predict(pruned_tree, testsub ,type = "prob")
#just want the probability of "Yes"
predictions = predictions[,1] 
#change response from factor to binary
test_binary = ifelse(testsub$suppdem=="Y", 1, 0)
d = -(1/nrow(testsub))*sum((test_binary *log(predictions) + (1-test_binary )*log(1-predictions)))
```

